Sketch Recognition and Understanding

### 2009-09-26 trunk/draft/System.tex
after approval of dr. samia

The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}. The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information. The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in pervious systems. The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols. This enables the system to be scalable for more complex shapes.

 %the performance of most algorithms is based on user style.

% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
%In stroke segmentation problem is Particle swarm optimization is used as the

\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm (PSO).  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.75]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{Block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines. The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive. This is an optimization problem, which evolutionary programming can solve efficiently. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve.

@@

 In \textbf{Feature Extraction} step, a feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

   The final step is \textbf{Classification} using a SVM classifier. The classifier attempts to classify the symbol into one the previously trained classes.% c  that will use the features computed to classify the segments into one of the previously trained classes. % Or to determine the symbol of the given segments from the set of preciously trained symbols.

 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\

%\chapter{System Details}
%\label{sec:SystemDetails}

\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maxima than the speed information. Agar et al. \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.  Similarly, direction information is used as it provides better distinctive maxima for the corners than the curvature information \cite{meanshift10}. Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

@@

The system uses a composite set of features which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\subsection{Structural and geometrical Features(FS1)}
Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these feature are computed. 
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.

		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order n\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These features represent the global shape of the symbol drawn.

 	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.
	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]
  \end{itemize}

@@

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier Paramters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the emprial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multiclass classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

 Both methods proved to be competitive with each other and there is no clear superiority of one method
over another \cite{Duan05whichis}. Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by expeirments.

@@

\section{Step by Step Example}
\label{secstepExample}
In this section a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two strokes and how the user draws them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process.

@@

\begin{figure}
	\centering
	\subfigure[First Stroke]{\label{fig:clockstroke1}
		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }

		\subfigure[Second Stroke]{	\label{fig:clockstroke2}
		\includegraphics[scale=0.7]{afterDefense/ClockComplete.jpg.eps}}

		\subfigure[Completed symbol]{
		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/Clock1.jpg.eps}	}
	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The arrows in the symbols illustrate how the user has drawn the symbol
	\label{fig:clock}
\end{figure}

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draw will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2} . Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the stroke points and the output is a list of segments that will be used in the recognition process. The algorithm that computes the segmentation is demonstrated as follows:
\begin{enumerate}

	\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing}
\begin{verbatim}
  1. For each point in Stroke do the following:
  		1.a Compute distance from previous point.
  		1.b Compute Time difference from previous point and add it to Time_list.  
  		1.c Compute Velocity and add it to Vel_list. 
  		1.d Compute Direction and add it to Dir_list.
  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
  	end (for loop)
  2. For each  list computed in previous step do the follwing:   
  	 2.a Compute average of all points. 
  	 2.b set average as threshold.  
  	 2.c Create region from continuous points larger than threshold. 
  	 2.d for each region do the following. 
  	 	   2.d.1 find the maximum 
  	 	   2.d.1 Add point to P_dp 
  	  end
     end 
\end{verbatim}	
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\begin{verbatim}
1. Initialize the axes of ellipse  and center by bounding box information
2. Repeat the following for few loops: 
	2.a Compute error in Equation  4.2 
	2.b Update center of ellipse and axis from error. 
3. Compute percent in Equation 4.4 
4. Computer eff in Equation 4.3 
5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
\end{verbatim}

\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm} 
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR Error Threshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness) 
         				3.a.1 Compute the Error of approximation as in Eq.4.5
         				3.a.2. Computer the fitness function as in Eq. 4.4 
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using Eq.3.2 
         3.e   Bound Velocity using Eq.3.4
         3.f   Update particle values using Eq. 3.3
         3.g   Refine the location of the particle 
         end .
     end 
4. return Gbest
\end{verbatim}	
\item Segmentation ALS2  as explained in Section \ref{sec:PolygonDivisionAlgorithm}
\begin{verbatim}
Initialize 
  1. Generate N particles with M dimensions (P_i...P_N) 
  2. Generate velocities for each particle for M dimension randomly. 
  3. Repeat till Maximum number of iteration OR Error Threshold  
         For each particle P_i do the following 
         3.a   Compute fitness  (fitness) 
              3.a.1 For each Segment in the particle do the following
         				    3.a.1.I   Error of approximation as Line (Eq. 4.7 - 4.9 )
         			      3.a.1.II  Error of approximation as circular arc (Eq. 4.10 - 4.19 )
         				    3.a.1.III Computer minimum of both approximation 
         			3.a.2 Compute Error of the particle as in Eq.  4.20
         			3.a.3 Compute fitness of the particle as in Eq. 4.21
         3.b   Determine Lbest  
               if (lbest_fitness<fitness)
                  3.b.1   lbest=P_i
         3.c      Determine Gbest.
               if (lGest_fitness<fitness)
                  3.c.1   lGest=P_i
         3.d   Compute the Velocity for each dimension using Eq.3.2 
         3.e   Bound Velocity using Eq.3.4
         3.f   Update particle values using Eq. 3.3
         3.g   Refine the location of the particle 
         end .
     end 
4. return Gbest
\end{verbatim}	
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{13cm}|}
\caption{Detailed Output of System in Each Step of Segmentation}
\label{tab:StepsStroke1} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{13cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\hline
\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{13cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

>>>>

The contribution in the system is mainly in attempting to generate the optimal segmentation using PSO which exhibited superiority with respect to similar algorithms in various other applications \cite{PolygonApproximationPSO}. The system attempts to find the optimal decomposition of the input stroke using PSO with the help of curvature and speed information. The use of PSO algorithm helps in eliminating the effects of input noise, problems of over and under segmentation that were in previous systems. The proposed symbol recognizer uses the generated segmentation to compute the feature vector which is used to identify more complex symbols. This enables the system to be scalable for more complex shapes.

%the performance of most algorithms is based on user style. 
% Stroke segmentation is one of the most complex problems in the sketch recognition system. It is dividing the   
%In stroke segmentation problem is Particle swarm optimization is used as the 
\section{System Overview}
\label{sec:AnOverviewOfTheSystem}
   This research solves sketch segmentation problem using particle swarm algorithm (PSO).  Due to sloppiness of users and hardware glitches the captured points must be processed to remove noise and redundancy before the system proceeds to compute the speed and curvature information. After that a set of possible corner or critical points is computed to guide the segmentation procedure.  A set of features is extracted from the segmented strokes to be used by the classifier.  %The stroke is the path of points from the instant the pen is down till it is up.
The block diagram in Figure \ref{fig:Blockdiagram} shows the main system blocks. The next section describes each block in details.
 %Due to sloppiness of users and hardware glitches the system processes the data before further
\begin{figure}[]
	\centering

\begin{center}
	\includegraphics[scale=0.9]{images/AllBlockDiagram.eps}
	\caption[The System Block Diagram]{Block diagram of the system}
	\label{fig:Blockdiagram}
\end{center}
\end{figure}

@@

The next step is \textbf{Segmentation}, the goal of the segmentation stage is to divide strokes into segments of either curves or lines. The segmentation problem is summarized as finding the best decomposition with least number of segments each represents a geometric primitive. This is an optimization problem, which evolutionary programming can solve efficiently. A genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al.\cite{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve. However, comparisons between PSO and GA \cite{ComparePSOGA05} shows that PSO converges to optimal solution faster than GA.

@@

 In \textbf{Feature Extraction} step, the approximated segments that were computed in the previous step are appended to a list of segments. A feature vector consists of a set of statistical, geometrical and spatial features are extracted from the segmented strokes. The feature vector is used as input for the next step.   %the system computes composite set of features some are statistical other are spatial features based on the type primitives.% The system extract segment based and statistical based features from the set of segments that the clustering algorithms produce.

   The final step is \textbf{Classification} using a SVM classifier. The classifier attempts to classify the symbol into one the previously trained classes.
   % c  that will use the features computed to classify the segments into one of the previously trained classes. % Or to determine the symbol of the given segments from the set of preciously trained symbols. 
 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\
%\chapter{System Details}
%\label{sec:SystemDetails}
\section{Possible Dominant Point Extraction}
\label{sec:Preprocessing}
%Some data is extracted from the points in the stroke.
The Possible Dominant Point Extraction step captures the points from the pointing devices then compute the information needed to determine a set of  \textit{Possible Dominant Points $P_{pd}$} which guides the segmentation step. It is noted as a person draws a shape the pen slows down near corners and picks up speed when drawing straight lines. Therefore, the speed information is widely used to identify the shape corners and edges \cite{earlyprocess}. The curvature information is used to determine the points with high angular changes along the path of points \cite{meanshift10}. Those observations help detecting the dominant points as the points with lower speeds and high curvatures.

Also the time difference between samples was used in \cite{polygonfeedback31} as it provides more distinct maximum than the speed information. Agar et al. \cite{polygonfeedback31} mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data. Figure \ref{fig:speed2Distance} shows the time difference and the speed graphs for the stroke drawn in Figure \ref{fig:orignalStroke}.  Similarly, direction information is used as it provides better distinctive maximum for the corners than the curvature information \cite{meanshift10}. Figure \ref{fig:curvatures} shows the direction and curvature graphs for the stroke drawn in Figure \ref{fig:orignalStroke}

@@

The system uses a composite set of features (total of 70 features) which include global shape properties like Rubine feature set \cite{gestureexample12},  ink density \cite{GeometryAndDomain102}, some appearance based properties like Zernike moments \cite{HeloiseBeautification}, and some stroke based structural information like number of perpendiculars lines , number of parallel lines and types of primitives in each symbol. The following list gives details on all the features used in the system.

\subsection{Structural and geometrical Features(FS1)}
 Features defines the structure of the geometrical symbol. Section \ref{sec:steprec} descripe in details how these 13 feature are computed.  
%\begin{description}
% \item [Structural and geometrical Features (FS1)] 
  \begin{itemize}
	 \item \emph{Segments:} Number of segments in the symbol.
	 \item \emph{Strokes:} Number of strokes or partial strokes that created the symbols.  
		\item  \emph{Primitives:} Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item \emph{Curves:} Number of curves or ellipses in the symbol. 
		\item \emph{Lines:} Number of lines in the symbol. 
		\item \emph{Perpendicular} lines Number of perpendicular lines. 
		\item \emph{Parallel} lines Number of parallel lines. 
		\item \emph{Intersections} Number of intersection between lines and curves. 
		\item \emph{T intersections} Number of T intersections. 
		\item \emph{L intersections} Number of L intersections. 
		\item \emph{X intersections} Number of X intersections.
		\item \emph{Min Radius} Minimum Radius of all curves in the symbol.
		\item \emph{Max Radius} Maximum Radius of all curves in the symbol.
		%\item[No. of holes] Number of holes in the symbols. 
\end{itemize}
%\begin{description}
%	 \item[Segments] Number of segments in the symbol.
%	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
%		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
%		\item [Curves] Number of curves or ellipses in the symbol. 
%		\item [Lines] Number of lines in the symbol. 
%		\item [Perpendicular lines] Number of perpendicular lines. 
%		\item [Parallel lines] Number of parallel lines. 
%		\item [Intersections] Number of intersection between lines and curves. 
%		\item [T intersections] Number of T intersections. 
%		\item [L intersections] Number of L intersections. 
%		\item [X intersections] Number of X intersections.
%		
%		
%		%\item[No. of holes] Number of holes in the symbols. 
%
%\end{description}
\subsection{Rubine Feature Set (FS2)} Features introduced by Rubine\cite{gestureexample12} for single stroke gestures. These 12 features represent the global shape of the symbol drawn. To compute the features we append all segments points into single path $ink_{path}$. The features is computed on the generated path $ink_{path}$.%then compute the features based on this path.  
\begin{enumerate}
	\item Cosine of starting angle.
	\item Sine of starting angle.
	\item Length of diagonal of bounding box. It gives an idea of the size of the bounding box).
	\item Angle of diagonal. It gives an idea of the shape of the bounding box (long, tall, square).
	\item Distance from start to end.  
	\item Total stroke length
	\item Change in Rotation( Arctan ). It gives the directional angle.
	\item Absolute rotation 
	\item Rotation squared 
	\item The maximum speed reached (squared) 
	\item Total time of stroke 
\end{enumerate}
\subsection{Statistical Features (FS3)} 
\begin{enumerate}
	\item \textsl{Zernike moments:} Zernike moments of order $n$\footnote{ We tried different values of n. Results show the analysis of each order.} \cite{HeloiseBeautification}.  For $n=10$ there are 32 moment features. 
\end{enumerate}
\subsection{Global shape properties set (FS4)}
 Features that are composites of different geometrical and statistical symbol. These 13 features represent the global shape of the symbol drawn.

 	\begin{itemize}
\item \emph{Size Ratio} Ratio between width to height of the symbol.
		\item \emp{Log of size Ratio} Log of the ratio between width and height of symbol.
	\item \emph{Ink density} Compute the density of points inside its bounding box\cite{GeometryAndDomain102}.   
 	\item \emph{Convex Hull Area} Area of convex hull with respect to area of bounding box of symbol.

	\item \emph{Convex Hull Perimeter} Perimeter of convex hull with respect to total length of symbol.
		\item \emph{Convex Hull Point} Number of points of convex hull with respect to Number of points of symbol.
		\item \emph{Mean Centroidal radius} The Mean of the centroidal radius which is the distance from each point in the symbol to the center of gravity.
				\item \emph{Center of Gravity} The center of gravity of the symbol (both in y and x direction).
					\item \emph{Mean Time} The mean of the time values between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]

	\item \emph{Mean Time difference} The mean of the time difference between each two successive points in the symbol. %Different strokes are appended to construct a single path.
	%\item [Ra]		
			\item \emp{Log of Area} Log of area of the symbol.
	\item \emph{Diff Area} Absolute difference of the area of bounding box and area of symbol.

  \end{itemize}

@@

%The foundations of Support Vector Machines (SVM) have
\subsubsection{SVM  Classification}
The resulting SVM method can be summarized as follows 
\begin{enumerate}
	\item Selecting the SVM classifier parameters. Parameter $C$ representing the tradeoff between minimizing the training set error and maximizing the margin. 
	\item Select A kernel function and their kernel parameters. For example for the radial basis function kernel, one must select width of the Gaussian $\gamma$. 
	\item Solve dual QP Equation \ref{eq:dualkernel} or an alternative SVM formulation using an appropriate quadratic programming or linear programming algorithm. 
	\item Get the support vector and recover the imperial threshold  variable $b$. 
	\item Classify new points using $f(x)=sign\left(\sum{i}y_i\alpha_i K\left(x,x_i\right)-b\right)$
\end{enumerate}
The parameters in step 1. are selected using cross validation if sufficient data are available. %However, recent model  selectin strategies can 
\subsubsection {Multi Class Classifiers}

 There is mainly two popular methods to change the binary SVM classification problem into Multi class classification problem. The first methods is One Versus All (OVA) Strategy where M binary classifiers is constructed one classifier for each class.  The $ith$ classifier is trained using positive example form $ith$ class and negative example from all other classes. The output function $p_i$ represents the distance and classification of the example from the $ith$ classifier. For a new example $x$, OVA SVM strategy assigns it to the class with the largest value of $p_i$.

@@

 Both methods proved to be competitive with each other and there is no clear superiority of one method
over another \cite{Duan05whichis}. Some claims that OVO strategy trains in less time than OVA strategy because they train with only portion of the dataset \cite{libsvm} but this claim have not yet been proofed by experiments.

@@

\section{Step by Step Example}
\label{secstepExample}
In this section a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two strokes and how the user draws them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process. 
\begin{figure}
	\centering
	\subfigure[First Stroke]{\label{fig:clockstroke1}
		\includegraphics[scale=0.7]{afterDefense/clockstroke1.jpg.eps} }
		\subfigure[Second Stroke] {	\label{fig:clockstroke2}
		\includegraphics[scale=0.7]{afterDefense/clockstroke2.jpg.eps}}		
		\subfigure[Completed symbol]{
		\label{fig:Clock1}\includegraphics[scale=0.7]{afterDefense/ClockDirections.jpg.eps}	}
	\caption{Stroke Drawing Sequence} a) The user first draws the ellipse b) The square wave is drawn inside the ellipse c) The final symbol the user draws along with arrows to illustrate how the directions the user used while drawing the symbol. 
	\label{fig:clock}
\end{figure}

@@

In this step the user draw one or more stroke to draw a symbol. Figure \ref{fig:clock} illustrate the sequence and direction of user draw. In this section, each stroke user draws will be labeled as it illustrated in Figure \ref{fig:clockstroke1} and Figure \ref{fig:clockstroke2}.  Table \ref{tab:StepsStroke1} shows the detailed output of the program at each step. The input of this step is the a set of strokes, each stroke has a set of  points and the output is a list of segments that will be used in the recognition process. More detailed example and algorithms are demonstrated in the Appendix \ref{ChapterstepExample} \\

For each stroke the user draw the system segments the stroke using the following procedure

\begin{enumerate}
\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}.
\item Ellipse Fitting: which is an attempt to fit as an ellipse as in details in Section \ref{sec:EllipseDetection} and Algorithm \ref{ellipseFitAlg} in Appendix \ref{ChapterstepExample}.
\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The algorithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}. (see Algorithm \ref{AlgS1alg} in Appendix \ref{ChapterstepExample})
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\item Segmentation ALS2 as explained in Section \ref{sec:PolygonDivisionAlgorithm}. ( see Algorithm \ref{AlgS2alg} in Appendix \ref{ChapterstepExample} ). 
\item Append best segmentation to List of segmentation ($L_s$) of this symbols. 
\end{enumerate}

\begin{landscape}
The steps of the first stroke (details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details}):
\begin{scriptsize}
%\begin{table}
	%\centering 
%\scalebox{0.6}{		
\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of First Stroke}
\label{tab:StepsStroke1} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\hline
\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

%Step & Discription of Step & Input & Steps & Output \\ \hline
Stroke 1 & Extracting point and computing boundary data & -- &   &

	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}

  \\ \hline
Possible Dominate point Extraction & for Computing details see Section\ref{sec:Preprocessing} &

	\includegraphics[scale=0.5]{afterDefense/clockstroke1.eps}
  &   &    
\\ \hline 
Segmentation & Segmentation of Ellipse & $P_{dp}$   &  Ellipse fitting  & 
The circle is   a = 103.05715108084719 ,b = 103.06851874747797 , Center( 283.03365107691775,  181.05035287667172 )   with Error =  1.9471676673301757 , Percent =1.0695193086531243
eff  = 0.27463462099275116

 \includegraphics[scale=0.5]{afterDefense/ellipse.jpg.eps}
 \\ \hline
 	\end{longtable}
 	\newpage
 	The steps of the second stroke:(details of the outputs are in  Appendix \ref{ChapterstepExample} in Table \ref{tab:StepsStroke1Details})

\begin{longtable}{|p{2cm}|p{2cm}|p{5cm}|p{2cm}|p{5cm}|}
\caption{ Output of System in Each Step of Segmentation of Second Stroke}
\label{tab:StepsStroke2} \\

\hline 
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
\multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
\multicolumn{1}{p{2cm}|}{\textbf{Details}} &
\multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endfirsthead

\multicolumn{3}{c}%
{{\bfseries \tablename\ \thetable{} -- continued from previous page}} \\ \hline
\multicolumn{1}{|p{2cm}|}{\textbf{Step}} & 
\multicolumn{1}{p{2cm}|}{\textbf{Description}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Inputs}} &
 \multicolumn{1}{p{2cm}|}{\textbf{Details}} &
 \multicolumn{1}{p{5cm}|}{\textbf{Output}} 
\\ \hline 
\endhead

Stroke 2 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} &
 	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps} \\ \hline
  & Possible Dominate point Extraction for details see Section\ref{sec:Preprocessing} &   	\includegraphics[scale=0.5]{afterDefense/clockstroke2.jpg.eps}  & 1. compute the $P_{pd}$  & 
    \\ \hline
